---
title: "Initial Korean covid 19"
author: "Leah Riganyan"
date: "04/01/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
The bar graph shows us that males are dying a lot more than females are. Looking at the entire population, women have a higher chance of being alive when compared to men.


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
   
The numerical results show us that 36% of females are dying, while the males are dying at 64%. Meaning that 56% of females are alive while only 44% of males are alive.

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results are showing us the amount of people that are dead and alive due to Covid. It shows us that 1831 females and 1411 males are alive. While 25 females and 44 males are dead. Looking at the observed count to the expected null, we can see that this goes against the expected null count. Covid seems to kill males over females, the odds ratio being 2.3.
   
   
   
## Conclusions
  
We took a look at the data sets for people in Korea and their rates of death compared by sex. Males seem to be dying more of Covid than females in Korea. The percentage difference is that 36% of women are dead, while 64% of men are dead. The sample looked at 1831 females and 1411 males, meaning that 25 women and 44 men died, showing us that Covid affects males a lot more compared to women.

